\sysfull

### 2018-08-05 paper/asplos19/background.tex
revisions to background (incomplete)
revisions to background (complete)

\textbf{What Hardware does \sys Target.} \sys is designed for the demands of existing and future energy-harvesting platforms based around general purpose, commodity computing components~\cite{wisp,msp430datasheet}. \sys targets a device with a
memory system that has fast, byte-addressable volatile and non-volatile memory; in particular, our target platform, WISP~\cite{wisp}, is equipped with a mixture of SRAM and FRAM. \sys leverages hardware support for fast, bulk-copying between memories via DMA~\cite{msp430datasheet}. \sys does not require architectural additions to commodity processors as in~\cite{su_date_2017,hicks_isca_2017,quickrecall,nvp}.

@@

\begin{figure}
    \begin{subfigure}{0.49\columnwidth}
        \includegraphics[width=\columnwidth]{figures/war-example.pdf}
        \caption{WAR-affected code}
        \label{fig:war-example}
    \end{subfigure}
    \begin{subfigure}{0.49\columnwidth}
        \includegraphics[width=\columnwidth]{figures/war-execution.pdf}
        \caption{Incorrect execution}
        \label{fig:war-execution}
    \end{subfigure}
    \caption{WAR dependency example. \texttt{NV} marks non-volatile variables, while \texttt{checkp()} is a checkpoint of volatile state. \task{replace checkp(); with checkpoint();}}
    \label{fig:war}
\end{figure}

@@

A power failure clears volatile state (e.g., registers and SRAM) while non-volatile memory (e.g., FRAM) persists. Upon a power failure, control flows to a prior point in the execution: by default, to the beginning of {\tt main()}. Early intermittent systems preserved progress by periodically checkpointing volatile execution context to non-volatile memory~\cite{mementos}, sometimes using hardware support~\cite{mottola2017harvos,hibernusplusplus,hibernus,idetic,quickrecall}. Checkpointing volatile state alone does not ensure data consistency when the system can directly manipulate non-volatile memory~\cite{mspcdino}. Precisely, data can get inconsistent when code includes a \emph{write-after-read} (WAR) dependency between operations that manipulate non-volatile memory. Fig.~\ref{fig:war} illustrates how state can become inconsistent in an intermittent execution using a simple example of an average operation over an array of integers.
The non-volatile variable \texttt{sum} introduces the WAR, being read and written sequentially by the increment operation. If a power failure occurs right before updating the non-volatile index \texttt{i} (\ref{fig:war-execution}, Line 7), \texttt{sum} gets erroneously incremented twice consecutively by the same array element (\ref{fig:war-execution}, Lines 6 and 9).

@@

\textbf{Frequent Non-volatile Memory Access.} 
\note{I want to emphasize again that we do not evaluate NV memory accesses}
Previous systems frequently access the non-volatile memory directly, which is a potential source of inefficiency.  Checkpointing systems copy all volatile
state~\cite{dino, mementos, ratchet, hicks_isca_2017} to non-volatile memory at
every checkpoint.  Task-based models, which do not take checkpoints, back up a
subset of non-volatile data to maintain memory consistency using, e.g.,
redo-logging.  Table~\ref{table:chechpoint_comparison} summarizes the data that
gets copied to and from the non-volatile memory in service of memory
consistency and progress preservation.  Ratchet~\cite{ratchet} and
Clank~\cite{hicks_isca_2017}, which require all memory to be non-volatile,
perform high-frequency non-volatile memory accesses, into checkpointing
frequently \note{is "into" a misspell?}.

\textbf{Incorrectly Sized Tasks Problem.} Compiler-inserted checkpoints or
programmer-defined tasks can be both non-terminating and/or
inefficient.  If a task (or code between two checkpoints) consumes more than the fixed,
maximum energy that the device can buffer, then the task
will never be able to complete using buffered energy only.  Such a task is
non-terminating, {\em prevents forward progress}, and makes the program
deadlock. 
%
If the task consumes far less energy than a device can buffer, the system
may operate {\em inefficiently}, saving the program state more often than needed.
%
Avoiding excessively costly, non-terminating, tasks and short, high-overhead,
tasks is challenging, because estimating the exact energy use of an arbitrary code is complicated.  
Moreover, when \emph{heterogeneous devices} with different energy buffers (e.g., 20 \si{\micro\farad}~\cite{rodriguez_tbcs_2015} to 0.1 \si{\farad}~\cite{moo}) are considered, the problem of porting a program becomes much worse
because a large task on one device may be a relatively short task
on a device with a larger buffer.
%
\textbf{\sys's execution model} solves this dilemma by introducing {\em task coalescing}, which
merges multiple tasks to amortize the overhead when tasks are too small, and
{\em task downscaling}, which breaks up a task when it is too large to complete.

>>>>

\textbf{What Hardware does \sys Target.} \sys is designed for the demands of existing and future \emph{tiny embedded energy-harvesting platforms} based around general purpose, commodity computing components~\cite{wisp,msp430datasheet}. \sys targets a device with a memory system that has fast, byte-addressable volatile and non-volatile memory; in particular, our target platform, WISP~\cite{wisp}, is equipped with a mixture of SRAM and FRAM. \sys leverages hardware support for fast, bulk-copying between memories via DMA~\cite{msp430datasheet}. \sys does not require architectural additions to commodity processors as in~\cite{su_date_2017,hicks_isca_2017,quickrecall,nvp}.

@@

\begin{figure}
    \begin{subfigure}{0.49\columnwidth}
        \includegraphics[width=\columnwidth]{figures/war-example.pdf}
        \caption{WAR-affected code}
        \label{fig:war-example}
    \end{subfigure}
    \begin{subfigure}{0.49\columnwidth}
        \includegraphics[width=\columnwidth]{figures/war-execution.pdf}
        \caption{WAR-affected code execution}
        \label{fig:war-execution}
    \end{subfigure}
    \caption{WAR dependency example. \texttt{NV} marks non-volatile variables, while \texttt{checkp()} is a checkpoint of volatile state. \note{replace checkp(); with checkpoint();} \note{reduce font size by one point}}
    \label{fig:war}
\end{figure}

@@

A power failure clears volatile state (e.g., registers and SRAM) while non-volatile memory (e.g., FRAM) persists. Upon a power failure, control flows to a prior point in the execution: by default, to the beginning of {\tt main()}. Early intermittent systems preserved progress by periodically checkpointing volatile execution context to non-volatile memory~\cite{mementos}, sometimes using hardware support~\cite{mottola2017harvos,hibernusplusplus,hibernus,idetic,quickrecall}. Checkpointing volatile state alone does not ensure data consistency when the system can directly manipulate non-volatile memory~\cite{mspcdino}. Precisely, data can get inconsistent when code includes a \emph{write-after-read} (WAR) dependency between operations that manipulate non-volatile memory. Fig.~\ref{fig:war} illustrates how state can become inconsistent in an intermittent execution using a simple example of an average operation over an array of integers.
The non-volatile variable \texttt{sum} introduces the WAR, being read and written sequentially by the increment operation. If a power failure occurs right before updating the non-volatile index \texttt{i} (\ref{fig:war-execution}, Line 7), \texttt{sum} gets erroneously incremented twice consecutively by the same array element (\ref{fig:war-execution}, Lines 6 and 9). \note{in the above example it should \ref{fig:war-example}, Line 6, first and then \ref{fig:war-execution}, Lines 5 and 8}

@@

\textbf{Frequent Non-volatile Memory Access.}  \note{I want to emphasize again that we do not evaluate NV memory accesses} Previous systems frequently access the non-volatile memory directly, which is a potential source of inefficiency. Checkpointing systems copy all volatile state~\cite{dino, mementos, ratchet, hicks_isca_2017} to non-volatile memory at every checkpoint.  Task-based models, which do not take checkpoints, back up a subset of non-volatile data to maintain memory consistency using, e.g., redo-logging.  Table~\ref{table:chechpoint_comparison} summarizes the data that gets copied to and from the non-volatile memory in service of memory consistency and progress preservation.  Ratchet~\cite{ratchet} and Clank~\cite{hicks_isca_2017}, which require all memory to be non-volatile, perform high-frequency non-volatile memory accesses, checkpointing frequently.

\textbf{Incorrectly Sized Tasks Problem.} Compiler-inserted checkpoints or programmer-defined tasks can be both non-terminating and/or inefficient.  If a task (or code between two checkpoints) consumes more than the fixed, maximum energy that the device can buffer, then the task will never be able to complete using buffered energy only.  Such a task is non-terminating, {\em prevents forward progress}, and makes the program deadlock. If the task consumes far less energy than a device can buffer, the system may operate {\em inefficiently}, saving the program state more often than needed. Avoiding excessively costly, non-terminating, tasks and short, high-overhead, tasks is challenging, because estimating the exact energy use of an arbitrary code is complicated. Moreover, when \emph{heterogeneous devices} with different energy buffers (e.g., 20 \si{\micro\farad}~\cite{rodriguez_tbcs_2015} to 0.1 \si{\farad}~\cite{moo}) are considered, the problem of porting a program becomes much worse because a large task on one device may be a relatively short task on a device with a larger buffer. \textbf{\sys's execution model} solves this dilemma by introducing {\em task coalescing}, which merges multiple tasks to amortize the overhead when tasks are too small, and {\em task downscaling}, which breaks up a task when it is too large to complete.
